---
title: "ease_aes - a walkthrough"
author: "Sarah Romanes"
date: "22 November 2018"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(gganimate)
library(tidyverse)
library(gridExtra)
```


# What does `ease_aes()` do?

> ease_aes() defines how different aesthetics should be eased during transitions.

I like to think of ease describing the *acceleration* of our transition. Should it have a constant speed? Should it speed up, or slow down? Or should it bounce around? We can control this using this component of `gganimate`.

There are **two** components to the `ease_aes` function, which are drawn from the `tweenr` package. They are 

* The name of the easing function to display: `ease`
* And the modifier for said `ease` function, in the form of `-in`, `-out`, or `-in-out`. 

Each ease_aes() command requires the ease function AND the modifier, *except* for the special case of the `linear function`. 


## The `ease` function

The ease function controls the *shape* of the acceleration. For example, `linear` will produce a constant speed for the transition, whereas `bounce` will provide a bouning effect! 

A list of the easing functions can be found [here](https://www.rdocumentation.org/packages/tweenr/versions/0.1.5/topics/display_ease), and they can easily be visualised with the use of the `tweenr` package as follows:

```{r}
par(mfrow=c(1,2))
tweenr::display_ease('linear')
tweenr::display_ease('bounce-in')
```


**Constant rate of change**

```{r}
ggplot(mtcars, aes(factor(cyl), mpg)) + 
  geom_boxplot() + 
  transition_states(
    gear,
    transition_length = 3,
    state_length = 1
  ) +
  enter_fade() + 
  exit_shrink() +
  ease_aes('linear')
```


**Bouncing**

```{r}
ggplot(mtcars, aes(factor(cyl), mpg)) + 
  geom_boxplot() + 
  transition_states(
    gear,
    transition_length = 3,
    state_length = 1
  ) +
  enter_fade() + 
  exit_shrink() +
  ease_aes('bounce-out')
```

# The affect of the modifiers

## Motivation

For the `linear` ease function, it is clear the rate of change is constant - so there is only one possible linear combination

```{r}
tweenr::display_ease('linear')
```

However, for the `quadratic` function, we have two options.

```{r}
par(mfrow=c(1,2))
tweenr::display_ease('quadratic-in')
tweenr::display_ease('quadratic-out')
```

It is clear that the role of the `-in` and `-out` modifiers is to change which *reflection* on the line *y=x* our function is on. This has the affect of changing concavity, or more pratically, the type of acceleration. In the `quadratic-in` example, our transition state is *speeding up* as time progresses. However, in `quadratic-out`, our transition state is *slowing down* as time progresses. 


**Speeding up**

```{r}
ggplot(mtcars, aes(factor(cyl), mpg)) + 
  geom_boxplot() + 
  transition_states(
    gear,
    transition_length = 3,
    state_length = 1
  ) +
  enter_fade() + 
  exit_shrink() +
  ease_aes('quadratic-in')
```


**Slowing down**

```{r}
ggplot(mtcars, aes(factor(cyl), mpg)) + 
  geom_boxplot() + 
  transition_states(
    gear,
    transition_length = 3,
    state_length = 1
  ) +
  enter_fade() + 
  exit_shrink() +
  ease_aes('quadratic-out')
```


### The `-in-out` modifier

The `in-out` modifier essentially appends the two accelerations together, speeding up and then slowing down. We can visualise this with the quadratic function as follows:

```{r}
tweenr::display_ease('quadratic-in-out')

ggplot(mtcars, aes(factor(cyl), mpg)) + 
  geom_boxplot() + 
  transition_states(
    gear,
    transition_length = 3,
    state_length = 1
  ) +
  enter_fade() + 
  exit_shrink() +
  ease_aes('quadratic-in-out')
```


# A more interesting example

We can use the `ggimage` package to add in the image of a car and visualise it transitioning from state (one side of the screen to the other) and changing the way it gets there using `ease_aes()`.

We first set up out static background -

```{r}
ggdat <- expand.grid(x=0.75 * c(0:14),
                     y=3* c(0:10),
                     fill=NA)


fill <- rep(NA, nrow(ggdat))

for(i in 1:nrow(ggdat)){
  if(ggdat$y[i]<=12){
    fill[i] <- "#009900"
  }
  else if(ggdat$y[i]==15){
    fill[i] <- "#72777f"
  } else
  {fill[i] <- "#4e86e0"}
}

ggdat$fill <- fill

ga <- ggplot(ggdat, aes(x, y)) +
  geom_tile(aes(fill=I(fill)), height=6, width=0.75, colour="white", size=1.2) +theme_void()
ga

```

And now we add a `bounce` function with `-in-out` modifier - creating the feeling of struggling to parallel park!

```{r}
dta <- data.frame(y =     c(15, 15),
                  x =     c(0,  10.5),
                  tstep = c("a" , "b"),
                  image = rep("images/car.png",2))


ga1 <- ga +geom_image(data=dta, aes(x,y, image=image), size=0.5)
ga1 + transition_states(tstep,
                     transition_length=1,
                     state_length = 1) +
  ease_aes('bounce-in-out')

```

