---
title: "Heart Pumping"
author: "Emi Tanaka"
date: "22/11/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This was working in the old version of `gganimate` but since upgrading to the new one, it has stopped working :(

```{r install pkgs, eval=FALSE}
## install from Github
devtools::install_github("jespermaag/gganatogram")

```

```{r}
library(gganimate)
library(gganatogram)
library(dplyr)
body <- hgMale_list$human_male_outline %>% 
  mutate(index=1:nrow(.)) %>%
  group_by(group) %>%
  summarise(x=NA, y=NA, index=0) %>%
  bind_rows(., hgMale_list$human_male_outline) %>% 
  arrange(group, index) %>% 
  select(x, y)
heart1 <- heart2 <- hgMale_list$heart[-1,]
heart2 <- heart2 %>% 
  mutate(x=1.3*(x - mean(x)) + mean(x),
         y=1.3*(y - mean(y)) + mean(y),
         state=2)
heart1 <- heart1 %>% 
  mutate(x=0.8*(x - mean(x)) + mean(x),
         y=0.8*(y - mean(y)) + mean(y),
         state=1)
heart <- rbind(heart1, heart2) %>% select(x, y, state)
```
I can't get this `transition_states` function to work and am not sure why.
It worked under an older version of gganimate. Then i updated the package and now I get an error?
this is the error

> Error in is_quosure(e2) : argument "e2" is missing, with no default

```{r transition_states, eval=FALSE}
ggplot(data=body, aes(x, -y)) +
  geom_path() +
  geom_polygon(fill="yellow", colour="black") +
  geom_path(data=heart, aes(x, -y)) + 
  geom_polygon(data=heart, aes(x, -y), fill="red") + 
  theme_void()  +
  ease_aes('quadratic-in-out') +
  
 #  transition_manual(state) 
  transition_states(state,
                    transition_length=3,
                    state_length=1) 

#animate(ga, nframes=10, width=350)  
```

BUt `transition_manual` works.  I also learnt that seeting the nframes argument of the `animate` function to a samll number makes the hear beat faster than a larger number

```{r transition_manual}
ga <- ggplot(data=body, aes(x, -y)) +
  geom_path() +
  geom_polygon(fill="yellow", colour="black") +
  geom_path(data=heart, aes(x, -y)) + 
  geom_polygon(data=heart, aes(x, -y), fill="red") + 
  theme_void()  +
  ease_aes('quadratic-in-out') +
  transition_manual(state) 
 

animate(ga, nframes=5, width=350)  
```

