---
title: "transition_manual"
author: "Adam Gruer + Saskia Freytag"
date: "22/11/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gganimate)
library(okcupiddata)
library(emo)
library(ggtextures)
theme_set(theme_minimal())
```

## Explain transistion_manual

A static plot
```{r static plot}
cars_plot <- ggplot(mtcars, aes(disp, mpg), ) + 
  geom_point(colour = "purple", size = 3) 
cars_plot

```

Facet by cylinder
```{r a facet}
cars_plot + facet_wrap(~cyl)

```

##Animate

Transition manual will show one frame per level of the supplied variable.
the `{current_frame}` can be used anywhere that accepts a string (I think!)
to display the value of the 'frame' variable at each step of the animation.

```{r animate}
cars_plot + transition_manual(cyl) +
  labs(title = "{current_frame}")
```

## More complex example

In this example we will be using the `okcupiddata` package to visualize how feelings about pets change with star sign. First let's begin with some data wrangling.   

```{r}
data(profiles)
profiles_red <- profiles %>% select(pets, sex, status, sign)

profiles_red <- profiles_red %>% mutate(sign = str_extract(sign, "[a-z']+[[:space:]]"))
profiles_red <- profiles_red %>% 
  mutate(pets_dislike = case_when(
    str_detect(pets, "dislikes dogs and dislikes cats") ~ "dogs_cats",
    str_detect(pets, "dislikes dogs") ~ "dogs", 
    str_detect(pets, "dislikes cats") ~ "cats"))
profiles_red <- profiles_red %>% mutate(sign = gsub(" ", "", sign))
profiles_red <- profiles_red %>% 
  mutate(pets_like = case_when(
    str_detect(pets, "likes dogs and likes cats") ~ "dogs_cats",
    str_detect(pets, "likes dogs") ~ "dogs", 
    str_detect(pets, "likes cats") ~ "cats"))
profiles_red <- profiles_red %>% 
  mutate(pets_has = case_when(
    str_detect(pets, "has dogs and has cats") ~ "dogs_cats",
    str_detect(pets, "has dogs") ~ "dogs", 
    str_detect(pets, "has cats") ~ "cats"))

# define factor
profiles_red <- profiles_red %>% filter(!is.na(sign))
profiles_red <- profiles_red %>% 
  mutate(sign = factor(sign, levels=c("aries", "taurus", "gemini", "cancer",
                                      "leo", "virgo", "libra", "scorpio",
                                      "sagittarius", "capricorn", "aquarius",
                                      "pisces"))) 
```

```{r}
profiles_red2 <- profiles_red %>% filter(!is.na(pets_dislike))
dislike_plot <- ggplot(profiles_red2, aes(x = sex, class=pets_dislike)) + geom_bar() 
dislike_plot + transition_manual(sign)

images = c(
 dogs= "https://www.akc.org/wp-content/themes/akc/component-library/assets//img/welcome.jpg",
 cats= "https://images.mentalfloss.com/sites/default/files/styles/mf_image_16x9/public/munchkinhed.png?itok=oeH4evcQ&resize=1100x1100",
 dogs_cats = "https://www.southernmallee.sa.gov.au/webdata/resources/images/Cat_and_Dog.jpg"
)

dislike_plot <- ggplot(profiles_red2, aes(pets_dislike, image = pets_dislike)) +
  geom_textured_bar() +
  scale_image_manual(values = images)

dislike_plot + transition_manual(sign) + labs(title="{current_frame}")
```

