---
title: "transition_components"
author: "Anna Quaglieri"
date: "22/11/2018"
output:
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = FALSE,cache=TRUE)
```

```{r}
library(gganimate)
library(tidyverse)
library(fivethirtyeight)
library(emo)
devtools::install_github("hadley/emo")
devtools::install_github("ropenscilabs/icon")
library(icon)
# List all the transitions
#ls("package:gganimate")
```


To understand how `transition_component` works I will use the `US_births_1994_2003` dataset from the `fivethirtyeight` package. The title of the article where this data was used is *Some People Are Too Superstitious To Have A Baby On Friday The 13th*. Is that true?

```{r}
head(US_births_1994_2003)
```

Some key points to keep in mind:

You need **id** and **time** components. The `transition_component` function is useful when you have the same subject (a plane, a day, a person, a neighborood etc..) with multiple observation over time. 

The first thing to keep in mind 

```{r eval=TRUE}
library(ggrepel)

fridays <- US_births_1994_2003 %>% 
  filter(day_of_week %in% c("Fri") & date_of_month %in% c(1,2,3,13,18,28))
table(fridays$date_of_month)

p=ggplot(fridays) + 
  geom_point(aes(x=year,y=births,colour=date_of_month)) +
  facet_wrap(~date_of_month)+
  transition_components(id=factor(date_of_month),time=date)+
  shadow_trail(distance = 0.01, size = 0.3)

animate(p, 200, 10,duration=20)
```

