---
title: "transition_filter"
author: "Anna Quaglieri"
date: "22/11/2018"
output:
  github_document:
    toc: true
---

```{r setup, include=FALSE,message=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache=TRUE)
```

```{r eval=FALSE,message=FALSE}
devtools::install_github("thomasp85/gganimate")
devtools::install_github("thomasp85/transformr")
```

```{r }
library(tidyverse)
library(gganimate)
library(transformr)
```


# `transition_components`

Let's start simple!! I believe that's where all big things start...

* `ggplot() + geom_point()`

```{r}
mydf <- mtcars
p1=ggplot(mydf, aes(x = hp, y = mpg)) +
geom_point()
p1
```
    
Let's add `transition_filter()`     
    
```{r}
p1 + transition_filter(transition_length = 10, 
                       filter_length = 0.2, 
                       wrap = TRUE, 
                       keep = FALSE,
                       qsec < 15,  qsec > 16) +
  labs(title="{closest_expression}")
```


## How do the filtering actually works?

`transition_filter` requires at least two logical conditions that it will use to produce the different instances of the animation. For example, in the example below `transition_filter` will plot first `hp` vs `mpg` only for rows that satisfies `qsec < 15` and then will plot `hp` vs `mpg` only for rows that satisfies `qsec > 16`. You can add as many conditions as you like!

Below I ran `transition_filter` with three filters and added `Label variables` to the title to give an idea what filters are used at each state.

```{r}
p1 + transition_filter(transition_length = 10, 
                       filter_length = 0.2, 
                       wrap = TRUE, 
                       keep = FALSE,
                       qsec < 14,  qsec > 16, qsec > 20) + 
  labs(title = "closest expression : {closest_expression}, \n transitioning : {transitioning}, \n closest_filter : {closest_filter}")
```

You can also provide a `TRUE\FALSE` column from your data frame as the logical condition!!!

```{r }
mydf <- mydf %>% mutate(cond1 = qsec < 15,
                        cond2 = qsec > 15 & qsec < 20,
                        cond3 = qsec > 20)

# update ggplot plo
p1=ggplot(mydf, aes(x = hp, y = mpg)) +
geom_point()

head(mydf[,c("cond1","cond2","cond3")])
```


```{r}
p1 + transition_filter(transition_length = 10, 
                       filter_length = 0.2, 
                       wrap = TRUE, 
                       keep = FALSE,
                       cond1,cond2) + 
  labs(title = "closest expression : {closest_expression}")
```

## Let's investigate the `keep` argument.
    
In the previous examples `keep` has always been `FALSE`. To make things a little bit exciting I will set it know to `FALSE`.  
    
```{r}
g <- p1 + transition_filter(transition_length = 10, 
                            filter_length = 0.2, 
                            wrap = TRUE, 
                            keep = TRUE,
                            qsec < 14,qsec > 16,qsec > 20) + 
  labs(title = "closest expression : {closest_expression}, \n transitioning : {transitioning}, \n closest_filter : {closest_filter}")

animate(g, nframes = 10, fps = 2)
```

Great! But... now I cannot really see what's being filtered at each stage. 

# Errors encountered along the way

If you ever get: `Error in transform_path(all_frames, next_filter, ease, params$transition_length[i],transformr is required to tween paths and lines` install the package `transformr`.

# Suggestion for improvement

* Maybe it would be good to add an option to have different colours for the observations that satisfy the regular expression. Something like a vector of length 1 or the same length as the number of conditions provided.


# Halloween Candy data!

Did you know how many types of Halloween Candies you could get?? I have been trying to understand how the sugar content and the price correlate depending whether is a `chocolate`, `fruity` or `caramel` candy! 

Not sure that makes much sense though :/ 

Probably, we should try to prevent the labels created from `ggrepel` from moving while transitioning, since the dots remains in the same position.
I have been trying using the `exclude_layer = 5` inside `transition_filter` that Danielle successfully used in `shadow_wake` https://github.com/ropenscilabs/learngganimate/blob/master/shadow_wake.md but it does not like it. 

```{r}
library(fivethirtyeight)
library(ggrepel)
data(candy_rankings)

p1=ggplot(candy_rankings) + 
  geom_point(aes(x=sugarpercent,y=pricepercent,colour=competitorname))+
  theme_bw()+
  theme(legend.position = "none") +
  geom_text_repel(aes(x=sugarpercent,y=pricepercent,colour=competitorname,label=competitorname)) +
  transition_filter(transition_length = 10, 
                            filter_length = 0.2, 
                            wrap = TRUE, 
                            keep = FALSE,
                    chocolate,fruity,caramel) + 
  labs(title = "closest expression : {closest_expression}")

animate(p1,nframes=50,duration=60)
```


```{r eval=FALSE,echo=FALSE}
library(fivethirtyeight)
library(ggrepel)
data(candy_rankings)
library(tweenr)
library(ggimage)

dta <- data.frame(y =     c(0.10, 0.10),
                  x =     c(0, 1),
                  tstep = c("a" , "b"),
                  image = rep("halloween_candies.jpg",2))


p1=ggplot(candy_rankings) + 
  geom_point(aes(x=sugarpercent,y=pricepercent,colour=competitorname))+
  theme_bw()+
  theme(legend.position = "none")

p2 <- p1 + geom_image(data=dta, aes(x,y, image=image), size=0.2)

p1=ggplot(candy_rankings) + 
  geom_point(aes(x=sugarpercent,y=pricepercent,colour=competitorname))+
  theme_bw()+
  theme(legend.position = "none") +
  geom_text_repel(aes(x=sugarpercent,y=pricepercent,colour=competitorname,label=competitorname)) +
  geom_image(data=dta, aes(x,y, image=image), size=0.2) + 
  transition_filter(transition_length = 1, 
                            filter_length = 0.2, 
                            wrap = TRUE, 
                            keep = FALSE,
                    chocolate,fruity,caramel) + 
  labs(title = "closest expression : {closest_expression}")

animate(p1,nframes=50,duration=60)
```