---
title: "transition_components"
author: "Anna Quaglieri"
date: "22/11/2018"
output:
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = FALSE,cache=TRUE)
```

```{r}
library(gganimate)
library(tidyverse)
library(fivethirtyeight)
library(emo)
devtools::install_github("hadley/emo")
devtools::install_github("ropenscilabs/icon")
library(icon)
# List all the transitions
#ls("package:gganimate")
```


To understand how `transition_component` works I will use the `US_births_1994_2003` dataset from the `fivethirtyeight` package. The title of the article where this data was used is *Some People Are Too Superstitious To Have A Baby On Friday The 13th*. Is that true?

```{r}
head(US_births_1994_2003)
```

Some key points to keep in mind:

You need **id** and **time** components. The `transition_component` function is useful when you have the same subject (a plane, a day, a person, a neighborood etc..) with multiple observation over time. 

The first thing to keep in mind 

```{r eval=TRUE}
library(ggrepel)

fridays <- US_births_1994_2003 %>% 
  filter(day_of_week %in% c("Fri") & date_of_month %in% c(1,2,3,13,18,28))
table(fridays$date_of_month)

p=ggplot(fridays) + 
  geom_point(aes(x=year,y=births,colour=date_of_month)) +
  facet_wrap(~date_of_month)+
  transition_components(id=factor(date_of_month),time=date)+
  shadow_trail(distance = 0.01, size = 0.3)

animate(p, 200, 10,duration=20)
```


# Example with US flights

Failed example!

```{r eval=FALSE, echo=FALSE}
library(fivethirtyeight)
library(nycflights13)
library(tidyverse)
library(sf)

# To plot the map I need:
# every flight is defined by a plane (tailnum) a departure airport
# link planes 
airports <- airports %>% rename(origin = faa)
infos_flights <- flights %>% left_join(planes) %>% left_join(airports) 

# Create topo background
earth <- st_as_sf(rnaturalearth::countries110)
lon_range <- range(infos_flights$lon) + c(-20, 20)
lat_range <- range(infos_flights$lat) + c(-20, 20)
bbox <- st_polygon(list(cbind(lon_range[c(1,1,2,2,1)], lat_range[c(1,2,2,1,1)])))
bbox <- st_sfc(bbox)
st_crs(bbox) <- st_crs(earth)
area <- st_intersection(earth, bbox)

infos_flights$id <- paste(infos_flights$flight,infos_flights$tailnum,
                          infos_flights$year,infos_flights$month,infos_flights$day,sep=".")
p <- ggplot(infos_flights[1:100,]) + 
  geom_sf(data = area, fill = 'white')+
  geom_point(aes(lon, lat, group = tailnum, colour = year), size = 2.5) +
  coord_sf(xlim = range(infos_flights$lon), ylim = range(infos_flights$lat))+
  
  theme(panel.background = element_rect(fill = 'lightblue'),
        legend.position = 'bottom') + 
  transition_components(tailnum, time_hour) + 
  shadow_trail(distance = 0.01, size = 0.3)

ggplot(infos_flights) + 
  geom_point(aes(x= ))

```


# Example with babynames

`shadow_trail` allows you to customise the way in which your observation leaves a trace of themself once they move on:

* `distance` let's you specify the distance between each trace left. I noticed that it does not work with a very small distance (0.001 wasn't working). It has something to do with the fact that `distance` is used a denominator at some steps and probably it gets too small
* `size` works like in the normal `ggplot()` (e.g. size of dots)

```{r}
library(devtools)
install_github("ropenscilabs/ozbabynames")
library(ozbabynames)
library(gganimate)

p=ggplot(ozbabynames[ozbabynames$name %in% c("Michael","James"),]) + 
  geom_point(aes(x=year,y=count,colour=name)) +
  theme_bw() + 
  transition_components(id=name,time=year)+
  shadow_trail(distance = 0.1, size = 2)
# leaves a trail after 0.1
p
```


```{r}
p=ggplot(ozbabynames[ozbabynames$name %in% c("Michael","James"),]) + 
  geom_point(aes(x=year,y=count,colour=name)) +
  transition_components(id=name,time=year)+
  shadow_trail(distance = 2, size = 2)
p
```

```{r eval=FALSE}
p=ggplot(ozbabynames[ozbabynames$name %in% c("Michael","James"),]) + 
  geom_point(aes(x=year,y=count,colour=name)) +
  transition_components(id=name,time=year)+
  shadow_trail(distance = 0.001, size = 2)
# Error in seq.default(1, params$nframes, by = params$distance) : 
# invalid '(to - from)/by'
# 
p
```

* Just an alternative with `transition_reveal`

```{r}
p2=ggplot(ozbabynames[ozbabynames$name %in% c("Michael","James"),]) + 
  geom_point(aes(x=year,y=count,colour=name,group=name)) +
  geom_line(aes(x=year,y=count,colour=name,group=name)) +
  geom_label_repel(aes(x=year,y=count,colour=name,label=name,group=name),
    arrow = arrow(length = unit(0.03, "npc"), type = "closed", ends = "first"),
    force = 10)+
  transition_reveal(id=name,along=year,keep_last = FALSE)+
  shadow_trail(distance = 0.01, size = 2,exclude_layer=3)
animate(p2,nframes = 100,duration = 30)
```

# What are the differences between `transition_reveal` and `transition_component`?

It look like they do sort of the same things... 

But!

* **geom_line()** with `transition_component()` throws and error: `Error: Unsupported layer type`

```{r eval=FALSE}
library(ggrepel)
p=ggplot(ozbabynames[ozbabynames$name %in% c("Michael","James"),]) + 
  geom_point(aes(x=year,y=count,colour=name)) +
  geom_line(aes(x=year,y=count,colour=name,group=name)) +
  transition_components(id=name,time=year)+
  shadow_trail(distance = 2, size = 2)
animate(p,nframes = 20)
```


```{r}
p2=ggplot(ozbabynames[ozbabynames$name %in% c("Michael","James"),]) + 
  geom_line(aes(x=year,y=count,colour=name,group=name)) +
  transition_reveal(id=name,along=year,keep_last = FALSE)+
  shadow_trail(distance = 0.01, size = 2)
animate(p2,nframes = 100,duration = 10)
```

* In order to show the transition across time you use `{frame_along}` from `transition_reveal` and `{frame_time}` in `transition_components`... 

```{r}
p1=ggplot(ozbabynames[ozbabynames$name %in% c("Michael","James"),]) + 
  geom_point(aes(x=year,y=count,colour=name)) +
  transition_reveal(id=name,along=year)+
  shadow_trail(distance = 0.01, size = 2)+
  labs(title="Year: {frame_along}")
p1
```

```{r}
p2=ggplot(ozbabynames[ozbabynames$name %in% c("Michael","James"),]) + 
  geom_point(aes(x=year,y=count,colour=name)) +
  transition_components(id=name,time=year)+
  shadow_trail(distance = 0.01, size = 2)+
  labs(title="Year: {frame_time}")
p2
```

