---
title: "transition_components"
author: "Anna Quaglieri"
date: "22/11/2018"
output:
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = FALSE,cache=TRUE)
```

```{r setup1, eval=FALSE}
devtools::install_github("hadley/emo")
devtools::install_github("ropenscilabs/icon")
```


```{r}
library(gganimate)
library(tidyverse)
library(fivethirtyeight)
```


# How does `transition_components()` work?

> ## Transition individual components through their own lifecycle
>
> This transition allows individual visual components to define their own "life-cycle". This means that the final animation will not have any common "state" and "transition" phase as any component can be moving or static at any point in time.

> ## Usage
> transition_components(id, time, range = NULL, enter_length = NULL,
  exit_length = NULL)


To understand how `transition_components()` works I will use the `US_births_1994_2003` dataset from the `fivethirtyeight` package. The title of the article where this data was used is *Some People Are Too Superstitious To Have A Baby On Friday The 13th*. Is that true?

```{r}
head(US_births_1994_2003)
```

Some key points to keep in mind:

In `transition_component()` you need and **id** and **time** components. The `transition_components()` function is useful when you have the same identifier (like a plane, a day, a person, a neighborood etc..) with multiple observation over time. 

In this example my **id** will be Fridays of the month. For example, one id if **Fri_1** which is the 1st Friday of a generic month across years. My plan is to compare the number of babies born on the Fridays 13th of different months (**Fri_13**) with baby born on other Fridays! To speed uo the process I will compare Fridays 13th with Fridays that occurs on the 1st, 2nd, 3rd, 18th and 28th across months and years.

```{r eval=TRUE}

fridays <- US_births_1994_2003 %>% 
  filter(day_of_week %in% c("Fri") & date_of_month %in% c(1,2,3,13,18,28))

p=ggplot(fridays) + 
  geom_point(aes(x=year,y=births,colour=date_of_month)) +
  facet_wrap(~date_of_month)+
  theme(legend.position = "bottom") +
  transition_components(id=factor(date_of_month),time=date)+
  shadow_trail(distance = 0.01, size = 0.3)

animate(p, nframes = 50, 10,duration=10)
```

To me, really, it doesn't seem like parents go out of their way to avoid births on Friday 13th!

# Example with babynames

The `babynames` is one of the great packages developed thanks to the effort of another #OzUnconf18 team https://github.com/ropenscilabs/ozbabynames! 

Below I am showing another example that uses `transition_components()` in combination with other fun animations like:

* `shadow_trail()` which allows you to customise the way in which your observations leave a trace of themselves once they move on with their transitions.

* Within `shadow_trail()`, the argument `distance` let's you specify the distance between each trace left. I noticed that it does not work with a very small distance (0.001 wasn't working). It has something to do with the fact that `distance` is used a denominator at some steps and probably it gets too small.
* The argument `size` works like in the normal `ggplot()` (e.g. size of dots) and it specify the size of trace left.

```{r}
# install_github("ropenscilabs/ozbabynames")
library(ozbabynames)

p=ggplot(ozbabynames[ozbabynames$name %in% c("Michael","James"),]) + 
  geom_point(aes(x=year,y=count,colour=name)) +
  theme_bw() + 
  transition_components(id=name,time=year)+
  shadow_trail(distance = 0.1, size = 2)
p
```

* Let's increase the `distance`

```{r}
p=ggplot(ozbabynames[ozbabynames$name %in% c("Michael","James"),]) + 
  geom_point(aes(x=year,y=count,colour=name)) +
  transition_components(id=name,time=year)+
  shadow_trail(distance = 2, size = 2)
p
```

* `distance` to small. The code below will throw the error:

> Error in seq.default(1, params$nframes, by = params$distance) : 
invalid '(to - from)/by'

```{r eval=FALSE}
p=ggplot(ozbabynames[ozbabynames$name %in% c("Michael","James"),]) + 
  geom_point(aes(x=year,y=count,colour=name)) +
  transition_components(id=name,time=year)+
  shadow_trail(distance = 0.001, size = 2)

p
```
