---
title: "transition_filter"
author: "Anna Quaglieri"
date: "22/11/2018"
output:
  github_document:
    toc: true
---

```{r setup, include=FALSE,message=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache=TRUE)
```

```{r eval=FALSE,message=FALSE}
devtools::install_github("thomasp85/gganimate")
devtools::install_github("thomasp85/transformr")
```

```{r }
library(tidyverse)
library(gganimate)
library(transformr)

# List all the transitions
#ls("package:gganimate")
```


# `transition_filter`

Let's start simple!! I believe that's where all big things start...

* `ggplot() + geom_point()`

```{r}
mydf <- mtcars
p1=ggplot(mydf, aes(x = hp, y = mpg)) +
geom_point()
p1
```
    
Let's add `transition_filter()`     
    
```{r}
p1 + transition_filter(transition_length = 10, 
                       filter_length = 0.2, 
                       wrap = TRUE, 
                       keep = FALSE,
                       qsec < 15,  qsec > 16) +
  labs(title="{closest_expression}")
```


## How do the filtering actually works?

`transition_filter` requires at least two logical conditions that it will use to produce the different instances of the animation. For example, in the example below `transition_filter` will plot first `hp` vs `mpg` only for rows that satisfies `qsec < 15` and then will plot `hp` vs `mpg` only for rows that satisfies `qsec > 16`. You can add as many conditions as you like!

Below I ran `transition_filter` with three filters and added `Label variables` to the title to give an idea what filters are used at each state.

```{r}
p1 + transition_filter(transition_length = 10, 
                       filter_length = 0.2, 
                       wrap = TRUE, 
                       keep = FALSE,
                       qsec < 14,  qsec > 16, qsec > 20) + 
  labs(title = "closest expression : {closest_expression}, \n transitioning : {transitioning}, \n closest_filter : {closest_filter}")
```

You cannot provide as the logical condition `TRUE\FALSE` columns from your data frame!.

```{r }
mydf <- mydf %>% mutate(cond1 = qsec < 15,
                        cond2 = qsec > 15 & qsec < 20,
                        cond3 = qsec > 20)
head(mydf[,c("cond1","cond2","cond3")])
```

The plot below will throw an error...

```{r eval=FALSE}
p1 + transition_filter(transition_length = 10, 
                       filter_length = 0.2, 
                       wrap = TRUE, 
                       keep = FALSE,
                       cond1,cond2) + 
  labs(title = "closest expression : {closest_expression}")
```

## Let's investigate the `keep` argument.
    
In the previous examples `keep` has always been `FALSE`. To make things a little bit exciting I will set it know to `FALSE`.  
    
```{r}
g <- p1 + transition_filter(transition_length = 10, 
                            filter_length = 0.2, 
                            wrap = TRUE, 
                            keep = TRUE,
                            qsec < 14,qsec > 16,qsec > 20) + 
  labs(title = "closest expression : {closest_expression}, \n transitioning : {transitioning}, \n closest_filter : {closest_filter}")

animate(g, nframes = 10, fps = 2)
```

Great! Well... now I cannot really see what's being filtered at each stage. 

# Errors encountered along the way

If you ever get: `Error in transform_path(all_frames, next_filter, ease, params$transition_length[i],transformr is required to tween paths and lines` install the package `transformr`.

# Suggestion for improvement

* Maybe it would be good to add an option to have different colours for the observations that satisfy the regular expression. Something like a vector of length 1 or the same length as the number of conditions provided.

